---
title: "assembly_stats"
author: "Allan Zuza"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# load packages and set functions
```{r load_packages, message=FALSE}
library(janitor)
library(kableExtra)
library(tidyverse)
library(data.table)
library(reshape2)
library(ggtree) 
library(phangorn)
library(ggnewscale)
library(patchwork)
library(ape)
library(rPinecone)
library(gridExtra)
'%notin%' <- Negate('%in%')
```

# plot attributes
```{r set_plot_atributes}
# set date colors

datecol <- c("Other" = "#f0f0f0", "2016" = "#cccccc",
             "2017" = "#969696", "2018" = "#636363",
             "2019" = "#252525" )

binnary_colors <- c("0" = "#f0f0f0", "1" = "#7570b3")

character_presence_absence <- presence <- c("no" = "#f0f0f0", "yes" = "#7570b3")

genotype_cols <- c("4.1.1" = "#EECC66", "4.3.1.1" = "#EE99AA",
                   "4.3.1.1.EA1" = "#6699CC", "4.3.1.2" = "#997700",
                   "4.3.1.2.EA2" = "#994455", "4.3.1.2.EA3" = "#004488")

yersinabactin_st_cols <-
  c("145-2LV" = "#EECC66", "151" = "#EE99AA", "277" = "#6699CC",
    "294" = "#997700", "327" = "#994455", "327-1LV" = "#004488", "78" = "#DDDDDD")

# rpinecone output
lineage_col <- c("0" = "#77AADD", "1" = "#EE8866")

sub_lineage_col <- c("7" = "#EEDD88", "8" = "#FFAABB", "Other" = "#99DDFF",
                    "6" = "#44BB99", "4" = "#BBCC33",  "#AAAA00", "#DDDDDD")


# saving directory
outdir <- "./results/plots/r/"
```


# load data frames for plot annotation
```{r load_data, message=FALSE}

# the Panaroo gene presence-absence matrix
data <- read.table("/Users/zuza/repos/redesigned-lamp/results/panaroo/gene_presence_absence.Rtab",
                   header = TRUE)
#head(data)

# load core gene tree from panaroo
tree <-
  read.tree("results/core_gene_tree/core_tree.treefile")


# load the snp based tree
snp_tree <-
  ape::read.tree("./results/iqtree/phylo.aln.treefile")

#load tree produced from gubbins
gubbins_tree <-
  ape::read.tree("./results/gubbins/ST39.final_tree.tre")

# read metadata from the Genome paper
tip_metadata <- 
  read_csv("docs/kleb_metadata.csv",
           show_col_types = FALSE) %>%
  clean_names() %>%
  filter(main_st == "ST39") %>%
  select(reference, strain, sample_acc, lane_acc, supplier_name,
         study_accession, supplier_sample_name, date, neonate, major_wards)

head(tip_metadata)

ariba <-
  read_csv("./docs/clean_ariba_amr.csv") %>%
  remove_constant()

plasmidfinder <-
  read_tsv("./docs/plamidfinder_summary.txt")
head(plasmidfinder)

# kleborate results
kleborate <-
  read_tsv("./results/kleborate/kleborate_results.txt")

kleborate
```


# lineage annotation files
```{r}
#install.packages("devtools")
#devtools::install_github("alexwailan/rpinecone")

results <- pinecone(gubbins_tree, 2, 3)

lineage_1 <- results$table %>% as.data.frame()

rownames(lineage_1) <- lineage_1$Taxa


lineage_1 <- lineage_1 %>% select(!c(Taxa))
lineage_1$sub_lineage <-
  case_when(grepl("singleton", lineage_1$`Sub-lineage`) ~ "Other",
                                   TRUE ~ lineage_1$`Sub-lineage`)


major_lineages <-
  lineage_1 %>%
  select(`Major.Sub-lineage`) %>%
  mutate(across(where(is.character), as.factor))


sub_lineages <-
  lineage_1 %>%
  select(sub_lineage) %>%
  mutate(across(where(is.character), as.factor))

summary(sub_lineages$sub_lineage)
```

# process data frames for plot annotation

```{r common_annotation_files}
# load date information from metadata file
date_data <- 
  tip_metadata %>%
  select(c(lane_acc, date, major_wards)) %>%
  filter(lane_acc %in% tree$tip.label)

date_data$year <-
  case_when(date_data$date == 2016 ~ "2016",
            date_data$date == 2017 ~ "2017",
            date_data$date == 2018 ~ "2018",
            date_data$date == 2019 ~ "2019",
            TRUE ~ "Other") 

date_data$wards <-
  case_when(date_data$major_wards ==
              "CHATINKHA NURSERY" ~ "CHATINKHA NURSERY",
            TRUE ~ "Other wards")

# again convert to a data.frame object
meta1 <- as.data.frame(date_data)

# assign rows
rownames(meta1) <- meta1$lane_acc

meta2 <- meta1 %>% select(year)

meta3 <- meta2 %>%
  mutate(year = as.factor(year))


head(date_data)

summary(as_factor(date_data$date))
summary(as_factor(date_data$major_wards))
summary(meta3$date)
```

## annotation for the core_gene tree
the dataframe from panaroo represents binary values as characters. there is a need
to convert them to numeric values then to binnary for optimal annotation
```{r create_heatmap_columns}

# filter the gene-presence-absense matrix to get genes in node 4196
df1 <-
  data %>% filter(Gene %in% c("rep_1", "xerC_1", "intA_3", "recF_1", "dam_2",
                              "xerC_4", "xerC_2", "pknD", "pknH")) 

# transpose the filtered matrix
transposed_df <- t(df1)

colnames(transposed_df) <-transposed_df[1,]

head(transposed_df)

# filter out the gene column
restructured_df <- transposed_df[rownames(transposed_df) %notin% c("Gene"), ]

# convert to a data.frame object because tibble handles rownames in a wierd way!
segment_genes <- as.data.frame(restructured_df)

head(rownames(segment_genes))

genes <-
  segment_genes %>%
  mutate(across(where(is.character), as.factor))


```

## annotations for the snp-based tree
```{r snp_based_tree_annotations}
amr_df <-
  ariba %>%
  as.data.frame() 

rownames(amr_df) <-
  amr_df$name

amr_df1 <- select(amr_df, !c(name))

head(amr_df1)

# plasmid data
plasmids <-
  plasmidfinder %>%
  as.data.frame()

# this column has extra characters which mess with parsing to subset the values
plasmids$`Col(pHAD28)_1` <-
  substr(plasmids$`Col(pHAD28)_1`, 1,4)

rownames(plasmids) <-
  plasmids$`#FILE`

plasmid_data <-
  select(plasmids, !c(`#FILE`, NUM_FOUND))

# replace all numeric to binary by matching the missing values
for (i in colnames(plasmid_data)) {
  plasmid_data[ ,i] <- if_else(plasmid_data[ ,i] %in% ".", 0, 1)
}

# plasmid_data_1 <-
#   plasmid_data %>%
#    mutate(across(where(is.character), as.integer))
# 
# for (i in colnames(plasmid_data_1)) {
#   plasmid_data_1[ ,i] <- if_else(!is.na(plasmid_data_1[ ,i]), 1, 0)
# }

plasmid_data_1 <-
  plasmid_data %>%
  mutate(across(where(is.numeric), as.factor))

```

## Kleborate output as annotations for phylogenetic trees
```{r}
unique_kleborate <-
  kleborate %>% remove_constant()


df1 <-
  unique_kleborate %>%
  select(strain, YbST, ybtS, ybtX, ybtQ, ybtA, irp2, irp1, ybtU, ybtT, ybtE, fyuA)

df1<- 
  df1 %>%
  mutate(across(where(is.character), as.factor))

df1[2:12] %>% summary()
df2 <- as.data.frame(df1)
rownames(df2) <- df2$strain

df3 <-
  df2 %>%
  select(!c(strain))

ybst <-
  df3 %>%
  select(YbST)


yb_genotype <-
  df3 %>%
  select(!c(YbST))



colnames(unique_kleborate)
unique_kleborate %>%
  mutate(across(where(is.double), as.factor)) %>%
  summary

unique_kleborate %>% group_by(iroD) %>% summary()

```


# plot the core gene trees
```{r plot_core_gene_tree, message=FALSE}
# root the tree at midpoint
tree <- midpoint(tree) # from library phangorn

y <-
  tree %>%
  ggtree() %<+%
  date_data +
    geom_tippoint(aes(color=wards), size = 1.5) +
     scale_color_viridis_d(na.translate = FALSE,
                        guide = guide_legend(order = 1))
y

h1 <- 
  gheatmap(y, meta3,
               width = 0.02,
         colnames_position = "top",
        colnames_angle = 90,
        colnames_offset_y = 3,
        hjust = 0.5) +
  scale_fill_manual(values = datecol,
                    na.translate = FALSE,
                        guide = guide_legend(order = 1)) +
  labs(fill = "Year")
  vexpand(0.01)

h1

h2 <- h1 + new_scale_fill()

h3 <- gheatmap(h2, genes,
         offset = 0.00004,
         width = 0.14,
           colnames_position = "top",
        colnames_angle = 90,
        colnames_offset_y = 4,
        hjust = 0.5) +
  scale_fill_manual(values = presence,
                       labels = c("Absent", "Present")) +
  labs(fill = "Gene Presence/Absence") +
  vexpand(0.02)

h3
```
```{r write_plot-TO-file, eval=FALSE}
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "core_gene_with_genes.png")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

png(plot_name,
    width = 800,
    height = 800)

h3

dev.off() 
```


# plot the snp based tree
## lineages
```{r plot_core_snp_tree_with_clane_data, message=FALSE}
#set local plot attributes
font_size <- 2

# root the tree at midpoint
tree_2 <- gubbins_tree
y <-
  tree_2 %>%
  ggtree() %<+%
  date_data +
    geom_tippoint(aes(color=wards), size = 1.5) +
     scale_color_viridis_d(na.translate = FALSE,
                        guide = guide_legend(order = 1)) +
  geom_treescale(label = "snps",
                 y = 130,
                 x  = 10) +
    geom_tiplab(align = TRUE,
                size = 1.5,
                linesize = 0.2) +
    hexpand(0.001)
y 

h1 <- 
  gheatmap(y, meta3,
               width = 0.01,
           offset = 100,
         colnames_position = "top",
        colnames_angle = 90,
        colnames_offset_y = 3,
        hjust = 0.5,
        font.size = font_size) +
  scale_fill_manual(values = datecol,
                    na.translate = FALSE,
                        guide = guide_legend(order = 1)) +
  labs(fill = "Year") +
  vexpand(0.025)

h1

h2 <- h1 + new_scale_fill()


lineage <- 
  gheatmap(h2, major_lineages,
         offset = 113,
         width = 0.01,
           colnames_position = "top",
        colnames_angle = 90,
        colnames_offset_y = 4,
        hjust = 0.5,
        font.size = font_size) +
  scale_fill_manual(values = lineage_col,
                       #labels = c("Absent", "Present"),
                    na.translate = FALSE) +
  labs(fill = "Major Sub lineages") +
  vexpand(0.02)

lineage
sublineage <- lineage + new_scale_fill()

sub_lineage <- 
  gheatmap(sublineage, sub_lineages,
         offset = 122,
         width = 0.01,
           colnames_position = "top",
        colnames_angle = 90,
        colnames_offset_y = 4,
        hjust = 0.5,
        font.size = font_size) +
  scale_fill_manual(values = sub_lineage_col,
                       #labels = c("Absent", "Present"),
                    na.translate = FALSE) +
  labs(fill = "Major Sub lineages") +
  vexpand(0.02)

sub_lineage

```

```{r write_plot-with-lineages-TO-file, eval=FALSE}
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "snp_based_with_lineage.png")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

png(plot_name,
    width = 800,
    height = 800)

sub_lineage

dev.off() 
```

## AMR data
```{r AMR_data}
y <-
  tree_2 %>%
  ggtree() %<+%
  date_data +
    geom_tippoint(aes(color=wards), size = 1.5) +
     scale_color_viridis_d(na.translate = FALSE,
                        guide = guide_legend(order = 1))
  
y 

h1 <- 
  gheatmap(y, meta3,
               width = 0.01,
         colnames_position = "top",
        colnames_angle = 90,
        colnames_offset_y = 3,
        hjust = 0.5,
        font.size = font_size) +
  scale_fill_manual(values = datecol,
                    na.translate = FALSE,
                        guide = guide_legend(order = 1)) +
  labs(fill = "Year") +
  vexpand(0.025)

h1
h2 <- h1 + new_scale_fill()

h3_amr <-
  gheatmap(h2, amr_df1,
         offset = 13,
         width = 0.4,
           colnames_position = "top",
        colnames_angle = 90,
        colnames_offset_y = 4,
        hjust = 0.5,
        font.size = font_size) +
  scale_fill_manual(values = character_presence_absence,
                       labels = c("Absent", "Present"),
                    na.translate = FALSE) +
  labs(fill = "Gene Presence/Absence") +
  vexpand(0.02)

h3_amr
```

```{r write_plot-with-AMR-TO-file, eval=FALSE}
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "snp_based_with_amr.png")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

png(plot_name,
    width = 800,
    height = 800)

h3_amr

dev.off() 
```

## plasmids
```{r Add_plasmid_data}
#set local plot attributes
font_size <- 2
angle <- 90

h1 <- 
  gheatmap(y, meta3,
               width = 0.01,
         colnames_position = "top",
        colnames_angle = angle,
        colnames_offset_y = 3,
        hjust = 0.5,
        font.size = font_size) +
  scale_fill_manual(values = datecol,
                    na.translate = FALSE,
                        guide = guide_legend(order = 1)) +
  labs(fill = "Year") +
  vexpand(0.025)

h1

h2 <- h1 + new_scale_fill()

h3_plasmid <- 
  gheatmap(h2, plasmid_data_1,
         offset = 13,
         width = 0.2,
           colnames_position = "top",
        colnames_angle = angle,
        colnames_offset_y = 4,
        hjust = 0.5,
        font.size = font_size) +
      scale_fill_manual(values = binnary_colors,
                    na.translate = FALSE,
                    labels = c("Absent", "Present"),
                    guide = guide_legend(order = 3)) +
  labs(fill = "Presence/Absence") +
  vexpand(0.02)

h3_plasmid
```

```{r write_plot-with-PLASMIDS-TO-file, eval=FALSE}
plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "snp_based_with_plasmids.png")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

png(plot_name,
    width = 800,
    height = 800)

h3_plasmid

dev.off() 
```

## Virulence factors

```{r virulence_factors_to_tree}
#set local plot attributes
font_size <- 2

# root the tree at midpoint
tree_2 <- midpoint(gubbins_tree) # from library phangorn

y <-
  tree_2 %>%
  ggtree() %<+%
  date_data +
    geom_tippoint(aes(color=wards), size = 1.5) +
     scale_color_viridis_d(na.translate = FALSE,
                        guide = guide_legend(order = 1)) 
y

h1 <- 
  gheatmap(y, meta3,
               width = 0.01,
         colnames_position = "top",
        colnames_angle = 90,
        colnames_offset_y = 3,
        hjust = 0.5,
        font.size = font_size) +
  scale_fill_manual(values = datecol,
                    na.translate = FALSE,
                        guide = guide_legend(order = 1)) +
  labs(fill = "Year") +
  vexpand(0.025)

h1

h2 <- h1 + new_scale_fill()

h3_yb <- 
  gheatmap(h2, ybst,
         offset = 13,
         width = 0.01,
           colnames_position = "top",
        colnames_angle = 90,
        colnames_offset_y = 4,
        hjust = 0.5,
        font.size = font_size) +
  scale_fill_manual(values = yersinabactin_st_cols,
                    na.translate = FALSE) +
  labs(fill = "ybST") +
  vexpand(0.02)

h3_yb
```


```{r collect_the_trees_and_save_them, eval=FALSE}
amr_plot <- 
  h3_amr + theme(legend.position = "none")

clade_plot <-
  sub_lineage + theme(legend.position = "none") 
  
plasmid_plot<-
  h3_plasmid + theme(legend.position = "none") + xlim(NA, 1265)

yb_plot<-
  h3_yb + theme(legend.position = "none") + xlim(NA, 1265)

grid.arrange(clade_plot, arrangeGrob(amr_plot, plasmid_plot, yb_plot), ncol = 2)


plotting_dir <- file.path(outdir, Sys.Date())
plot_name <- file.path(plotting_dir, "snp_based_trees.pdf")

if (!dir.exists(plotting_dir)) {dir.create(plotting_dir, recursive = TRUE)}

pdf(plot_name,
    width = 32,
    height = 18,
    pointsize = 24)

grid.arrange(clade_plot, arrangeGrob(amr_plot, plasmid_plot, yb_plot), ncol = 2)

dev.off() 

```







# miscellanious

```{r, eval=FALSE}
kleb_metadata <-
  read_csv("/Users/zuza/repos/redesigned-lamp/docs/kleb_metadata.csv")
date_data <-
  select(kleb_metadata, LaneAcc, DATE)


y <- transpose(date_data, make.names = "LaneAcc")

merge(data, y)
head(data)

head(date_data)
date_data <- column_to_rownames(date_data, var = "LaneAcc")
head(date_data)




merged_data <- merge(transposed_df, date_data) 
head(merged_data)

data.table:transpose(merged_data)

x <- reshape2::melt(merged_data)

ggplot(x, aes(y = Gene, x = variable, fill = value)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 90))

```


```{r, eval=FALSE}

pdf("~/repos/redesigned-lamp/test1.pdf", width = 300, height = 20)
pivot %>% ggplot(aes(Gene, count)) + geom_tile(color = "red") +
  theme(axis.text.x = element_text(angle = 90))

dev.off()
```


### attribute results from Graphia
Using Graphia, I extract the attribute data after clustering the nodes using the
Louvain Clusterring algorithym.
```{r load_graphia_results, message=FALSE, eval=FALSE}
library(tidyverse)
clusters <- 
  read_csv("./results/graphia/full_final_graph-attributes.csv") %>% clean_names()

# save objects to Rdata
save(clusters,
     file = "data/r/pangenomeClusters.RData")
```
```{r}
load("data/r/pangenomeClusters.RData")
clusters

summary(clusters)
str(clusters)

clusters %>% filter(louvain_cluster_size <= 118)
clusters %>%
  filter(node_members != 134) %>%
  filter(!str_detect(node_name_2, "group")) %>%
  #group_by(louvain_cluster) %>%
  ggplot() +
  geom_col(aes(node_members, x = node_name_2)) +
  theme(axis.text.x = element_text(angle = 90))


clusters %>%
  filter(node_members != 134) %>%
  filter(node_members >= 100) %>%
  filter(!str_detect(node_name_2, "group")) %>%
  #group_by(louvain_cluster) %>%
  ggplot() +
  geom_col(aes(node_members, x = node_name_2)) +
  theme(axis.text.x = element_text(angle = 90))

genes_to_check <- clusters %>%
  filter(node_members < 134) %>%
  #filter(node_members >= 100) %>%
  filter(!str_detect(node_name_2, "group"))

#genes_to_check$node_name_2
```


### check genes from graphia output
```{r eval=FALSE}
# check genes from graphia output
genes <- genes_to_check$node_name_2
genes


df1 <- data %>% filter(str_detect(Gene, paste(genes, collapse = "|"))) %>%
  filter(!str_detect(Gene, "~"))

transposed_df <- t(df1)

colnames(transposed_df) <-transposed_df[1,]

head(transposed_df)

restructured_df <- transposed_df[rownames(transposed_df) %notin% c("Gene"), ]

x <- as.data.frame(restructured_df)

#rownames(x)

tree <- midpoint(tree)
y <- tree %>% ggtree() %<+% x

gheatmap(y, x,
              # width = 0.02,
         colnames_position = "top",
        colnames_angle = 90,
        colnames_offset_y = 14,
        hjust = 0.5) +
  scale_fill_viridis_d(direction = -1) +
  vexpand(0.1)
```

